---
title: "Methods"
output: pdf_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tinytex)
library(tidyverse)
library(arsenal)

theme_set(theme_minimal())
```

# Data Cleaning

```{r message = FALSE, warning = FALSE}
hate_df =
  read_csv("./data/HateCrimes.csv") %>% 
  mutate(
    state = as.factor(state),
    unemployment = as.factor(unemployment),
    urbanization = as.factor(urbanization),
    hate_crimes_per_100k_splc = as.numeric(hate_crimes_per_100k_splc)
  )
```

# Descriptive Statistics

```{r}
# Table labels
my_labels =
  list(
    unemployment = "Antibody IgM",
    urbanization = "Urbanization",
    median_household_income = "Median Household Income",
    perc_population_with_high_school_degree = "Percent with HS Degree",
    perc_non_citizen = "Percent Non-Citizen",
    gini_index = "Gini Index",
    perc_non_white = "Percent Non-White",
    hate_crimes_per_100k_splc = "Hate Crimes per 100k"
)

# Table controls
my_controls = tableby.control(
  total = F,
  test = F,
  numeric.stats = c("N", "meansd", "medianq1q3", "range", "Nmiss2"),
  cat.stats = c("N", "countpct"),
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max",
    Nmiss2 = "Missing",
    countpct = "N (%)",
    N = "N"
    )
  )

# Generate table
descriptive_tab =
  tableby( ~ unemployment + 
           urbanization +
           median_household_income +
           perc_population_with_high_school_degree +
           perc_non_citizen +
           gini_index +
           perc_non_white +
           hate_crimes_per_100k_splc,
           data = hate_df,
           control = my_controls)

summary(
  descriptive_tab,
  title = "Descriptive Statistics: Hate Crimes Data",
  labelTranslations = my_labels,
  text = T)
```

As a note, I didn't include the "states" variable as the output was huge and not that helpful. Suggest we include a note somewhere that data from 50 states + Washington, DC.

# Distribution of Outcome Data

Histogram of raw outcome data (hate crimes per 100k).

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(x = hate_crimes_per_100k_splc, y = ..density..)) +
  geom_histogram(bins = 11) +
  geom_density(alpha = 0.2, color = "blue") +
  labs(
    x = "Hate Crimes per 100k",
    y = "Density"
  )
```

These data look skewed :(

Histogram of log-transformed outcome data (hate crimes per 100k).

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(x = log(hate_crimes_per_100k_splc), y = ..density..)) +
  geom_histogram(bins = 11) +
  geom_density(alpha = 0.2, color = "blue") +
  labs(
    x = "Hate Crimes per 100k (Log Scale)",
    y = "Density"
  )
```

Looks better!

Box plot of the (raw) outcome data.

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(y = hate_crimes_per_100k_splc)) +
  geom_boxplot() +
  labs(
    y = "Hate Crimes per 100k"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
    )
```

Just based on the boxplot, it ks like there are two states with potential usually high rates (Washington, DC and Oregon).

# Examining Potential Multicollinearity

```{r}
hate_df %>% 
  select(
    hate_crimes_per_100k_splc,
    median_household_income,
    perc_population_with_high_school_degree,
    perc_non_citizen,
    gini_index,
    perc_non_white
    ) %>%
  cor(use = "complete.obs") %>% # Ignoring NA values
  round(., 2)
```

Based on this output, the following pairs of variables have a correlation of 60% or higher:

* Percentage non-citizens & percentage non-white (0.75)

* Median household income & percentage of population with a high school degree (0.65)

```{r}
hate_df %>% 
  select(
    hate_crimes_per_100k_splc,
    median_household_income,
    perc_population_with_high_school_degree,
    perc_non_citizen,
    gini_index,
    perc_non_white
    ) %>% 
  pairs()
```

# Simple Linear Regression Using Income Inequality (Per FiveThirtyEight)

Fitting SLR using income inequality (measured by Gini index) per FiveThirtyEight findings.

```{r}
slr_income_lm = lm(hate_crimes_per_100k_splc ~ gini_index, data = hate_df)
slr_income_log_lm = lm(log(hate_crimes_per_100k_splc) ~ gini_index, data = hate_df)

summary(slr_income_lm)
summary(slr_income_log_lm)
```

Gini index appears to be a significant predictor only when using the raw outcome data (not the log-transformed outcome data).

# Jacy's Ideas

```{r eval = FALSE}
##Project ideas
hate = read.csv("/Users/jacysparks/Downloads/HateCrimes.csv")
head(hate)
dim(hate)
hate$hate_crimes_per_100k_splc = as.character(hate$hate_crimes_per_100k_splc)
hate$hate_crimes_per_100k_splc = as.numeric(hate$hate_crimes_per_100k_splc)
summary(hate)
##Four NA's for outcome
##NA for Wyoming, South Dakota, North Dakota, and Idaho
hate[,c(1,9)]
##Could remove 
hate = na.omit(hate)

##3 NA's for non citizen


##Create indicators
names(hate)[names(hate)=="unemployment"] = "High.Unemployment"
names(hate)[names(hate)=="urbanization"] = "High.Urban"
names(hate)[names(hate)=="median_household_income"] = "Med.Income"
names(hate)[names(hate)=="perc_population_with_high_school_degree"] = "HS.Degree"
names(hate)[names(hate)=="perc_non_citizen"] = "Non.Citizen"
names(hate)[names(hate)=="perc_non_white"] = "Non.White"
names(hate)[names(hate)=="hate_crimes_per_100k_splc"] = "Hate.Crime"
hate$High.Unemployment = ifelse(hate$High.Unemployment=="high",1,0)
hate$High.Urban = ifelse(hate$High.Urban=="high",1,0)

##Outcome var is skewed
hate$Hate.Crime = log(hate$Hate.Crime)
hist(hate$Hate.Crime)
##Much better

reg = lm(Hate.Crime~.-state,data=hate)
summary(reg)

pairs(hate[,4:9],lower.panel=NULL)
cor(hate[,4:9])
#Percent white and percent non-white highly correlated


##Check linearity
for(i in 4:8){
  plot(hate[,i],hate$Hate.Crime,main=colnames(hate)[i])
}
plot(reg)

```

