---
title: "Methods"
output: pdf_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tinytex)
library(tidyverse)
library(arsenal)

theme_set(theme_minimal())
```

# Data Cleaning

```{r message = FALSE, warning = FALSE}
hate_df =
  read_csv("./data/HateCrimes.csv") %>% 
  mutate(
    state = as.factor(state),
    unemployment = as.factor(unemployment),
    urbanization = as.factor(urbanization),
    hate_crimes_per_100k_splc = as.numeric(hate_crimes_per_100k_splc)
  )
```

# Descriptive Statistics

```{r}
# Table labels
my_labels =
  list(
    unemployment = "Unemployment",
    urbanization = "Urbanization",
    median_household_income = "Median Household Income",
    perc_population_with_high_school_degree = "Percent with HS Degree",
    perc_non_citizen = "Percent Non-Citizen",
    gini_index = "Gini Index",
    perc_non_white = "Percent Non-White",
    hate_crimes_per_100k_splc = "Hate Crimes per 100k"
)

# Table controls
my_controls = tableby.control(
  total = F,
  test = F,
  numeric.stats = c("N", "meansd", "medianq1q3", "range", "Nmiss2"),
  cat.stats = c("N", "countpct"),
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max",
    Nmiss2 = "Missing",
    countpct = "N (%)",
    N = "N"
    )
  )

# Generate table
descriptive_tab =
  tableby( ~ unemployment + 
           urbanization +
           median_household_income +
           perc_population_with_high_school_degree +
           perc_non_citizen +
           gini_index +
           perc_non_white +
           hate_crimes_per_100k_splc,
           data = hate_df,
           control = my_controls)

summary(
  descriptive_tab,
  title = "Descriptive Statistics: Hate Crimes Data",
  labelTranslations = my_labels,
  text = T)
```

As a note, I didn't include the "states" variable as the output was huge and not that helpful. Suggest we include a note somewhere that data from 50 states + Washington, DC.

# Distribution of Outcome Data

Histogram of raw outcome data (hate crimes per 100k).

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(x = hate_crimes_per_100k_splc, y = ..density..)) +
  geom_histogram(bins = 11) +
  geom_density(alpha = 0.2, color = "blue") +
  labs(
    x = "Hate Crimes per 100k",
    y = "Density"
  )
```

These data look skewed :(

Histogram of log-transformed outcome data (hate crimes per 100k).

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(x = log(hate_crimes_per_100k_splc), y = ..density..)) +
  geom_histogram(bins = 11) +
  geom_density(alpha = 0.2, color = "blue") +
  labs(
    x = "Hate Crimes per 100k (Log Scale)",
    y = "Density"
  )
```

Looks better!

Box plot of the (raw) outcome data.

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(y = hate_crimes_per_100k_splc)) +
  geom_boxplot() +
  labs(
    y = "Hate Crimes per 100k"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
    )
```

Just based on the box plot, it looks like there are two states with potential usually high rates (Washington, DC and Oregon).

# Examining Potential Multicollinearity

```{r}
hate_df %>% 
  select(
    hate_crimes_per_100k_splc,
    median_household_income,
    perc_population_with_high_school_degree,
    perc_non_citizen,
    gini_index,
    perc_non_white
    ) %>%
  cor(use = "complete.obs") %>% # Ignoring NA values
  round(., 2)
```

Based on this output, the following pairs of variables have a correlation of 60% or higher:

* Percentage non-citizens & percentage non-white (0.75)

* Median household income & percentage of population with a high school degree (0.65)

```{r}
hate_df %>% 
  select(
    hate_crimes_per_100k_splc,
    median_household_income,
    perc_population_with_high_school_degree,
    perc_non_citizen,
    gini_index,
    perc_non_white
    ) %>% 
  pairs()
```

# Simple Linear Regression Using Income Inequality (Per FiveThirtyEight)

Fitting SLR using income inequality (measured by Gini index) per FiveThirtyEight findings.

```{r}
slr_gini_lm = lm(hate_crimes_per_100k_splc ~ gini_index, data = hate_df)
slr_gini_log_lm = lm(log(hate_crimes_per_100k_splc) ~ gini_index, data = hate_df)

summary(slr_gini_lm)
summary(slr_gini_log_lm)
```

Gini index appears to be a significant predictor only when using the raw outcome data (not the log-transformed outcome data).

Scatter plots associated with these simple linear regression models.

```{r message = FALSE, warning = FALSE}
hate_df %>% 
  ggplot(aes(x = gini_index, y = hate_crimes_per_100k_splc)) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(
    x = "Gini Index",
    y = "Hate Crimes per 100k"
    ) +
  geom_abline(intercept = -1.5275, slope = 4.0205)

hate_df %>% 
  ggplot(aes(x = gini_index, y = log(hate_crimes_per_100k_splc))) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(
    x = "Gini Index",
    y = "Hate Crimes per 100k (Log Scale)"
    ) +
  geom_abline(intercept = -3.676, slope = 4.932)
```

Diagnostic plots of these two simple linear regression models.

```{r}
par(mfrow = c(2, 2))
plot(slr_gini_lm)
plot(slr_gini_log_lm)
```

Normality looks better when we perform the log transformation on the outcome data. However, for both versions of this model, we can see an outlying value in the upper right corner (this corresponds to Washington, DC).

# Trying Stepwise Approach

First, looking at the full model (with and without log transformation of outcome).

```{r}
hate_nona_df = # Removing rows with missing values from the dataset
  hate_df %>% 
  drop_na()

full_lm = lm(
  hate_crimes_per_100k_splc
  ~ unemployment + 
    urbanization +
    median_household_income +
    perc_population_with_high_school_degree +
    perc_non_citizen +
    gini_index +
    perc_non_white,
  data = hate_nona_df
)

full_log_lm = lm(
  log(hate_crimes_per_100k_splc)
  ~ unemployment + 
    urbanization +
    median_household_income +
    perc_population_with_high_school_degree +
    perc_non_citizen +
    gini_index +
    perc_non_white,
  data = hate_nona_df
)

summary(full_lm)

summary(full_log_lm)
```

Trying stepwise approach.

```{r}
step(full_lm, direction = "both") # Trying "both" directions

step(full_log_lm, direction = "both") # Trying "both" directions
```

This procedure is retaining the following two predictors:

* Precent population with high school degree

* Gini index

# Jacy's Ideas

```{r eval = FALSE}
##Project ideas
hate = read.csv("/Users/jacysparks/Downloads/HateCrimes.csv")
head(hate)
dim(hate)
hate$hate_crimes_per_100k_splc = as.character(hate$hate_crimes_per_100k_splc)
hate$hate_crimes_per_100k_splc = as.numeric(hate$hate_crimes_per_100k_splc)
summary(hate)
##Four NA's for outcome
##NA for Wyoming, South Dakota, North Dakota, and Idaho
hate[,c(1,9)]
##Could remove 
hate = na.omit(hate)

##3 NA's for non citizen


##Create indicators
names(hate)[names(hate)=="unemployment"] = "High.Unemployment"
names(hate)[names(hate)=="urbanization"] = "High.Urban"
names(hate)[names(hate)=="median_household_income"] = "Med.Income"
names(hate)[names(hate)=="perc_population_with_high_school_degree"] = "HS.Degree"
names(hate)[names(hate)=="perc_non_citizen"] = "Non.Citizen"
names(hate)[names(hate)=="perc_non_white"] = "Non.White"
names(hate)[names(hate)=="hate_crimes_per_100k_splc"] = "Hate.Crime"
hate$High.Unemployment = ifelse(hate$High.Unemployment=="high",1,0)
hate$High.Urban = ifelse(hate$High.Urban=="high",1,0)

##Outcome var is skewed
hate$Hate.Crime = log(hate$Hate.Crime)
hist(hate$Hate.Crime)
##Much better

reg = lm(Hate.Crime~.-state,data=hate)
summary(reg)

pairs(hate[,4:9],lower.panel=NULL)
cor(hate[,4:9])
#Percent white and percent non-white highly correlated


##Check linearity
for(i in 4:8){
  plot(hate[,i],hate$Hate.Crime,main=colnames(hate)[i])
}
plot(reg)

```


Fit model based on stepwise (log transformed outcome data) and make a scatterplot with regression line, regenerate diagnostic plots using this model
```{r stepwise}
stepwise_log_lm = lm(
  log(hate_crimes_per_100k_splc)
  ~ perc_population_with_high_school_degree + # Question - which is the main predictor we're putting on the x axis?
    gini_index,
  data = hate_nona_df)

hate_nona_df %>% 
  ggplot(aes(x = perc_population_with_high_school_degree, y = log(hate_crimes_per_100k_splc))) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(
    x = "Percentage of Population with High School Degree",
    y = "Hate Crimes per 100k (Log Scale)"
    ) +
  geom_abline(intercept = -18.947, slope = 11.554)
  # ab line not showing up...

hate_nona_df %>% 
  ggplot(aes(x = gini_index, y = log(hate_crimes_per_100k_splc))) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(
    x = "Gini Index",
    y = "Hate Crimes per 100k (Log Scale)"
    ) +
  geom_abline(intercept = -18.947, slope = 16.486) # ab line not showing up...

par(mfrow = c(2, 2))
plot(stepwise_log_lm)
```

Formalize if DC is an influential point quantitatively using the Cook’s value

Using Cook's D and studentized residuals, DC is not a cause for concern. Cook's D is 0.332, not high enough for concern. The studentized residual is not greater than 2.5 for any variable. However, DFFIT is greater than 1. Could be cause for concern. Regression analysis shows it is not influential, so no need to delete.
```{r}
plot(full_log_lm)
influence.measures(full_log_lm)
stu_res<-rstandard(full_log_lm)
outliers_y<-stu_res[abs(stu_res)>2.5]
outliers_y

summary(full_log_lm)
full_nodc = lm(log(hate_crimes_per_100k_splc)~.-state,data=hate_nona_df)
summary(full_nodc)
```

Use Box-Cox method
```{r}
MASS::boxcox(full_lm) # default grid of lambdas is -2 to 2 by 0.1

# Could change grid of lambda values just to zoom in, get more precise
MASS::boxcox(full_lm, lambda = seq(-1, 1, by=0.05) ) 


## i'll just QQ plot these as she does in lecture

qqnorm(resid(full_lm), xlab = "Expected Value", ylab = "Residual", main = "")
qqline(resid(full_lm))
title("(a) QQ Plot for Y (Hate Crimes per 100k)")

qqnorm(resid(full_log_lm), xlab = "Expected Value", ylab = "Residual", main = "")
qqline(resid(full_log_lm))
title("(d) QQ Plot lnY (Ln(Hate Crimes per 100k)")

#Note that this is not strictly better. It's up for interpretation. 
```
The optimal value of Y^a is near 0, indicating that a natural log transformation of the outcome  for all practical intents and purposes is optimal. We proceed with the log transformation. 


Check for interactions

Fit a linear model with all two-way interactions and investigated the interactions which were significant. There were two where there was a significant interaction between a categorical and continuous variable, between median household income and unemployment, and median household income and urbanization. Looking at the interaction plots between these variables we do see that the lines for the two levels of the categorical variables cross, indicating interaction. Stratifying on unemployment we see that the coefficient for median household income is positive when unemployment is low and negative when unemployment is high. When we stratify on urbanization, we see that the coefficient for median household income has a higher magnitude when urbanization is low (although both coefficients are negative). These stratified analyses indicate that interactions do exist.
```{r}
full_int_lm = lm(log(hate_crimes_per_100k_splc)~(.-state)^2, data = hate_nona_df)
summary(full_int_lm)

hate_nona_df %>% 
  ggplot(aes(x = median_household_income, y = log(hate_crimes_per_100k_splc), color = unemployment)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Median Household Income",
       y = "log(Hate Crimes Per 100k splc)",
       title = "Interaction Plot for Unemployment and Median Household Income")

hate_nona_df %>% 
  ggplot(aes(x = median_household_income, y = log(hate_crimes_per_100k_splc), color = urbanization)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Median Household Income",
       y = "log(Hate Crimes Per 100k splc)",
       title = "Interaction Plot for Urbanization and Median Household Income")

# Stratified analysis for unemployment

unemployment_low = hate_nona_df %>% 
  filter(unemployment == "low") %>% 
  select(-unemployment)

unemployment_high = hate_nona_df %>% 
  filter(unemployment == "high") %>% 
  select(-unemployment)

unemployment_low_lm = lm(log(hate_crimes_per_100k_splc)~.-state, data = unemployment_low)
summary(unemployment_low_lm)

unemployment_high_lm = lm(log(hate_crimes_per_100k_splc)~.-state, data = unemployment_high)
summary(unemployment_high_lm)

## Median household income coefficients have different signs when unemployment is low vs high

urbanization_low = hate_nona_df %>% 
  filter(urbanization == "low") %>% 
  select(-urbanization)

urbanization_high = hate_nona_df %>% 
  filter(urbanization == "high") %>% 
  select(-urbanization)

# Stratified analysis for urbanization

urbanization_low_lm = lm(log(hate_crimes_per_100k_splc)~.-state, data = urbanization_low)
summary(urbanization_low_lm)

urbanization_high_lm = lm(log(hate_crimes_per_100k_splc)~.-state, data = urbanization_high)
summary(urbanization_high_lm)

## Median household income coefficients have different magnitudes when urbanization is low vs high
```

I tried including these interaction terms in the full model with gini index and percent of population with high school degree; none of these interactions were significant so they were omitted in the full model.
```{r}
interaction_log_lm = lm(
  log(hate_crimes_per_100k_splc)
  ~ perc_population_with_high_school_degree + gini_index + median_household_income + unemployment + urbanization + median_household_income*unemployment + median_household_income*urbanization,
  data = hate_nona_df)

summary(interaction_log_lm)
```



